---
title: "My Dashbord at DACSS"
bibliography: testVisual.bib
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
---

```{r setup, include=FALSE}
library(flexdashboard) # always
library(ggplot2)
library(dplyr)
library(scales)
```

# HW1
## Intro {.sidebar}

Preparing the dashboard based on the instructions given at @magallanes_reyes_data_2022, and @sievert_flexdashboard_2022.

### References

::: {#refs}
:::

## Column {data-width="450"}

### Worst barplot <!-- a title for plot -->

```{r}
##for del1Draft.rds
link1="https://github.com/98oakOrg/690V_w1/raw/main/del1Draft.rds"
readRDS(url(link1))
```

### Worst histogram

```{r}
## for del2Draft.rds
link2="https://github.com/98oakOrg/690V_w1/raw/main/del2Draft.rds"
readRDS(url(link2))
```

## Column {data-width="550"}

### Worst scatterplot

```{r}
## for del3Draft.rds
link3="https://github.com/98oakOrg/690V_w1/raw/main/del3Draft.rds"
readRDS(url(link3))
```


# HW2: mtcars {data-orientation=rows}

## Row {data-height=500}

### Column {data-width=6}
#### Count of Cars per GEAR
```{r, fig.width=5, fig.height=3}
gear_counts <- mtcars %>%
  count(gear) %>%
  mutate(gear = factor(gear, levels = sort(unique(gear), decreasing = TRUE)))

ggplot(gear_counts, aes(x = gear, y = n)) +
  geom_col(fill = "grey80") +
  geom_text(aes(label = n), color = "blue", hjust = -0.15, size = 7) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Count of Cars per GEAR",
    subtitle = "THREE-GEAR CARS ARE DOMINANT",
    x = NULL,
    y = "Gear",
    caption = "Source: mtcars data"
  ) +
  theme_minimal(base_size = 14)
```

### Column {data-width=6}
#### Count of Cars per CARB

```{r, fig.width=5, fig.height=3}
carb_counts <- mtcars %>%
  count(carb) %>%
  arrange(carb)

ggplot(carb_counts, aes(x = factor(carb), y = n)) +
  geom_col(fill = "yellow") +
  geom_text(aes(label = n), vjust = -0.4, color = "darkgreen", size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Count of Cars per CARB",
    subtitle = "FEW CARS have SIX or more CARBURATORS",
    x = "CARB",
    y = NULL,
    caption = "Source: mtcars data"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold"))
```

***

## Row {data-height=720}
### Column {data-width=12}
#### Distribution of Cylinder Counts by Carburetors and Gears

```{r, fig.width=12, fig.height=6}
dist_df <- mtcars %>%
  count(gear, carb, cyl) %>%
  mutate(
    pct = n / nrow(mtcars),
    pct_label = percent(pct, accuracy = 0.1)
  )

ggplot(dist_df, aes(x = factor(cyl), y = n)) +
  geom_col(fill = "yellow") +
  geom_text(aes(label = pct_label), vjust = -0.4, size = 3) +
  facet_grid(
    gear ~ carb,
    switch = "y",
    labeller = labeller(
      carb = function(x) paste0("carb: ", x),
      gear = function(x) paste0("gear: ", x)
    )
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Distribution of Cylinder Counts by Carburetors and Gears",
    x = "cylinders",
    y = "count",
    caption = "Source: mtcars data"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey80", color = NA)
  )
```