---
title: "My Dashbord at DACSS"
bibliography: testVisual.bib
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
---

```{r setup, include=FALSE}
library(flexdashboard) # always
library(ggplot2)
library(dplyr)
library(scales)
library(readxl)
library(forcats)
library(reshape2)
library(patchwork)
```

# HW1
## Intro {.sidebar}

Preparing the dashboard based on the instructions given at @magallanes_reyes_data_2022, and @sievert_flexdashboard_2022.

### References

::: {#refs}
:::

## Column {data-width="450"}

### Worst barplot <!-- a title for plot -->

```{r}
##for del1Draft.rds
link1="https://github.com/98oakOrg/690V_w1/raw/main/del1Draft.rds"
readRDS(url(link1))
```

### Worst histogram

```{r}
## for del2Draft.rds
link2="https://github.com/98oakOrg/690V_w1/raw/main/del2Draft.rds"
readRDS(url(link2))
```

## Column {data-width="550"}

### Worst scatterplot

```{r}
## for del3Draft.rds
link3="https://github.com/98oakOrg/690V_w1/raw/main/del3Draft.rds"
readRDS(url(link3))
```


# HW2: mtcars {data-orientation=rows}

## Row {data-height=500}

### THREE-GEAR CARS ARE DOMINANT {data-width=6}

```{r, fig.width=5, fig.height=3}
gear_counts <- mtcars %>%
  count(gear) %>%
  mutate(gear = factor(gear, levels = sort(unique(gear), decreasing = FALSE)))

ggplot(gear_counts, aes(x = gear, y = n)) +
  geom_col(fill = "grey80") +
  geom_text(aes(label = n), color = "blue", hjust = -0.15, size = 7) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Count of Cars per GEAR",
    x = "Count",
    y = "Gear",
    caption = "Source: mtcars data"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.title.x = element_blank()
  )
```

### FEW CARS have SIX or more CARBURATORS {data-width=6}
#### Count of Cars per CARB

```{r, fig.width=5, fig.height=3}
carb_counts <- mtcars %>%
  count(carb) %>%
  arrange(carb)

ggplot(carb_counts, aes(x = factor(carb), y = n)) +
  geom_col(fill = "yellow") +
  geom_text(aes(label = n), vjust = -0.4, color = "darkgreen", size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Count of Cars per CARB",
    x = "CARB",
    y = NULL,
    caption = "Source: mtcars data"
  ) +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(face = "bold"))  +
  theme(
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.y = element_blank()
  )
```

***

## Row {data-height=720}
### Only 5-gear cars utilize the 6 and 8 carburetor setups. {data-width=12}


```{r, fig.width=12, fig.height=6}
dist_df <- mtcars %>%
  count(gear, carb, cyl) %>%
  mutate(
    pct = n / nrow(mtcars),
    pct_label = percent(pct, accuracy = 0.1)
  )

ggplot(dist_df, aes(x = factor(cyl), y = n)) +
  geom_col(fill = "yellow") +
  geom_text(aes(label = pct_label), vjust = -0.4, size = 3) +
  facet_grid(
    gear ~ carb,
    labeller = labeller(
      carb = function(x) paste0("carb: ", x),
      gear = function(x) paste0("gear: ", x)
    )
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Distribution of Cylinder Counts by Carburetors and Gears",
    x = "cylinders",
    y = "count",
    caption = "Source: mtcars data"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey80", color = NA)
  )
```

# HW3 Univariate

## Column

### Categorical Variable: School Type

```{r}
#read in file and set data needed for chart

mass_data <- read_excel("ncesdata_MASS_202425 (1).xlsx", sheet = "dataMASS", skip = 11)


mass_data <- mass_data %>%
  mutate(
    Students = as.numeric(Students),
    Teachers = as.numeric(Teachers),
    `Student Teacher Ratio` = as.numeric(ifelse(`Student Teacher Ratio` %in% c("†", "–", "‡"), NA, `Student Teacher Ratio`)),
    `Directly Certified` = as.numeric(ifelse(`Directly Certified` %in% c("†", "–", "‡"), NA, `Directly Certified`))
  )

#plot1 categorical
type_table <- as.data.frame(table(mass_data$Type, useNA = "ifany"))
names(type_table) <- c("School_Type", "Count")
type_table$Percent <- prop.table(type_table$Count) * 100

ggplot(type_table, aes(x = reorder(School_Type, -Percent), y = Percent)) +
  geom_col(fill = "gray") +
  geom_text(aes(label = paste0(round(Percent, 1), "%")), hjust = 0, size = 3) +
  coord_flip() + 
  labs(
    title = "Types of Public Schools in Massachusetts",
    caption = "Source: NCES CCD 2024-2025 public school data",
    x = NULL,
    y = "Percent"
  ) +
  scale_y_continuous(labels = label_number(suffix = "%"),
  expand = expansion(mult = c(0, 0.2))
  ) +
  theme(
    plot.caption = element_text(
      size = 8,
      face = "italic",
      margin = margin(t = 8)
    )
  )
  
```

### Numerical Variable: Students

```{r}
#plot 2 numerical
mean_students <- mean(mass_data$Students, na.rm = TRUE)

ggplot(mass_data, aes(x = Students)) +
  geom_histogram(binwidth = 100, fill = "black") +
  geom_vline(xintercept = mean_students, color = "red", linewidth = 1) +
  annotate(
    "text",
    x = mean_students + 350,
    y = 300,
    label = paste0("Mean = ", round(mean_students, 0)),
    color = "red",
    vjust = -0.5,
    size = 4
  ) +
  labs(
    title = "Student Enrollment Distribution",
    caption = "Source: NCES CCD 2024-2025 public school data",
    x = "Number of Students Enrolled",
    y = "Count of Schools"
  ) +
  theme(
    plot.caption = element_text(
      size = 8,
      face = "italic",
      margin = margin(t = 8)
    )
  )
```


# HW4: Multivariate {data-orientation=rows}
## Row

### Covid: Cases vs Deaths
```{r}
covid <- read.csv("covid-19.csv")
#names(covid)

#summary(covid$daily_cases_mean)
#summary(covid$daily_deaths_mean)

covid_nozero <- covid %>%
  filter(daily_cases_mean > 0, daily_deaths_mean > 0)

library(ggplot2)

ggplot(covid_nozero, aes(x = daily_cases_mean, y = daily_deaths_mean)) +
  geom_point(alpha = 0.3) +
  scale_x_log10(labels = scales::label_number(accuracy = 1)) +
  scale_y_log10(labels = scales::label_number(accuracy = 1)) +
  labs(
    title = "Relationship Between Average Daily Cases and Deaths",
    x = "Average Daily Cases (log scale)",
    y = "Average Daily Deaths (log scale)") +
  theme_minimal()
```

## Row
### UCR Code by Sex

```{r}

arrests <- read_excel("Mass_Arrests.xlsx", sheet = 1)
#invisible(names(arrests))
#invisible(glimpse(arrests))


#determine distribution of data
library(dplyr)
charge_distribution <- arrests %>%
  count(`Arrest Offense by UCR Code`, sort = TRUE) %>%
  mutate(pct = n / sum(n),
    cum_pct = cumsum(pct))
#charge_distribution

#determine what I'll report on
top_charges <- charge_distribution %>%
  filter(cum_pct <= 0.80) %>% 
  pull(`Arrest Offense by UCR Code`)

arrests_categorize <- arrests %>%
  mutate(Charge_grouped = ifelse(
      `Arrest Offense by UCR Code` %in% top_charges,
      `Arrest Offense by UCR Code`,
      "Other"),
    Charge_label = case_when(
      Charge_grouped == "90D" ~ "Drug Possession",
      Charge_grouped == "90Z" ~ "Drug Violation",
      Charge_grouped == "MV"  ~ "Motor Vehicle",
      TRUE ~ "Other"))


arrests_categorize$Charge_label <- fct_infreq(arrests_categorize$Charge_label)

ggplot(arrests_categorize, aes(x = Sex, fill = Charge_label)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Distribution of UCR Code by Sex",
    x = "Sex",
    y = "Proportion",
    fill = "Offense Type") +
  theme_minimal()
```

### Offense Type by Age

```{r}
arrests_numcat <- arrests_categorize %>%
  filter(Charge_label != "Other")

ggplot(arrests_numcat, aes(x = Charge_label, y = Age)) +
  geom_boxplot(fill = "grey80") +
  coord_flip() +
  labs(
    title = "Distribution of Offense Types by Age",
    x = "Offense Type",
    y = "Age") +
  theme_minimal()
```

## Row
### Multivariate Improvement: Safe Cities



```{r}
# load file
safe <- read_excel("safeCities_ecoUI.xlsx")
safe <- safe[, !(names(safe) %in% "Overall_Score")]

# long format
safeL <- melt(safe, id.vars = "city")

# create dimensions (first letter mapping)
safeL$dimension <- substr(safeL$variable, 1, 1)

# average into dimensions
safe_avg <- safeL %>%
  group_by(city, dimension) %>%
  summarize(value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    dimension = recode(
      dimension,
      "D" = "Governance",
      "H" = "Health",
      "I" = "Infrastructure",
      "P" = "Personal",
      "E" = "Environment"))

# lower volume to top 15 cities (based on overall average)
top_cities <- safe_avg %>%
  group_by(city) %>%
  summarize(avg = mean(value), .groups = "drop") %>%
  arrange(desc(avg)) %>%
  slice(1:15) %>%
  pull(city)

# filter both datasets
safeL_small <- safeL %>%
  filter(city %in% top_cities)

safe_avg_small <- safe_avg %>%
  filter(city %in% top_cities)

#  city order for both plots
safeL_small$city <- factor(safeL_small$city, levels = rev(top_cities))
safe_avg_small$city <- factor(safe_avg_small$city, levels = rev(top_cities))

# original (full heatmap)
viz1 <- ggplot(safeL_small, aes(x = variable, y = city, fill = value)) +
  geom_tile() +
  labs(title = "Original: All Variables") +
  theme_minimal() +
  theme(axis.text.x = element_blank())

# improved (simplified dimension heatmap)
viz2 <- ggplot(safe_avg_small, aes(x = dimension, y = city, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Improved: Averaged Dimensions",
    x = "Dimension",
    y = "",
    fill = "Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8))

# show viz side-by-side
(viz1 | viz2) + plot_layout(widths = c(1.3, 1))
```

